---
title: "Thermal Behavior Technical Report"
author: "John Stanton-Geddes"
date: '2014-04-17'
output: pdf_document
---

# Introduction

# Methods

# Results

Analysis of foraging behavior by heat-shocked and control ants in experimental even-aged colonies of *Pogonomyrmex barbatus*. 

```{r libraries}
library(pander)
library(stringr)
library(ggplot2)
```


```{r load_data}
data <- read.csv("data/CCThermBe_observations.csv", stringsAsFactors = FALSE)
str(data)
```

Five levels of 'forager_color' when we only expected 4. Check this:

```{r}
unique(levels(data$forager_color))
```

Okay - some "red " with trailing space. Change these to "red" by removing trailing spaces from all levels.

```{r}
data$forager_color <- str_trim(data$forager_color)
unique(data$forager_color)
```

Assign treatment to colors for each colony.

```{r treatment}
# if colony == 1 & color == blue trt <- control
test <- vector(length = nrow(data))

data$trt <- ifelse(data$colony == 1 & data$forager_color == "blue", "control", 
                    ifelse(data$colony == 1 & data$forager_color == "red", "heat", 
                           ifelse(data$colony == 2 & data$forager_color == "blue", "control", 
                                  ifelse(data$colony == 2 & data$forager_color == "green", "heat", 
                                         NA))))
# check
table(data$forager_color, data$trt, data$colony)
```

For each colony, chi-squared test to evaluate if the proportion of ants foraging depends on heat treatment

```{r colony1}
data$trt <- as.factor(data$trt)
str(data)

data1 <- data[which(data$colony == 1 & !is.na(data$trt)), ]
str(data1)  

chisq.test(as.numeric(data1$trt))
qplot(data1$trt, xlab = "Treatment", ylab = "Number of foragers")
```

No difference in foraging among treatments - almost exactly 50:50.

Same analyis for colony 2.

```{r colony2}
data2 <- data[which(data$colony == 2 & !is.na(data$trt)), ]
str(data2)  

chisq.test(as.numeric(data2$trt))
qplot(data2$trt, xlab = "Treatment", ylab = "Number of foragers")
```


Possible that the effects of heat-shock on foraging behavior diminished over time as many of the workers died in the colonies over time. So we next try the analysis focusing on the first two weeks after treatment only.

```{r early_only_c1}
# for colony 1, extract first three sampling dates
# convert date info into format 'mm/dd/yyyy'
data1$date <- as.Date(data1$date, "%m/%d/%Y")
str(data1)

# select first 4 observation dates
data1.early <- data1[which(data1$date %in% unique(data1$date)[1:4]), ]
str(data1.early)

chisq.test(as.numeric(data1.early$trt))
qplot(data1.early$trt, xlab = "Treatment", ylab = "Number of foragers")
```


Still no difference for colony 1. In fact, more of the foragers were from the "control" group.

Test for colony 2.

```{r early_only_c2}
# convert date info into format 'mm/dd/yyyy'
data2$date <- as.Date(data2$date, "%m/%d/%Y")
str(data2)

# select first 4 observation dates
data2.early <- data2[which(data2$date %in% unique(data2$date)[1:4]), ]
str(data2.early)

chisq.test(as.numeric(data2.early$trt))
qplot(data2.early$trt, xlab = "Treatment", ylab = "Number of foragers")
```

Same as for colony 1. 