---
title: "Thermal Behavior Technical Report"
author: "John Stanton-Geddes"
date: '2014-04-17'
output: html_document
---

Analysis of foraging behavior by heat-shocked and control ants in experimental even-aged colonies of *Pogonomyrmex barbatus*. 

```{r libraries}
library(pander)
library(stringr)
```


```{r load_data}
data <- read.csv("data/CCThermBe_observations.csv")
str(data)
```

Five levels of 'forager_color' when we only expected 4. Check this:

```{r}
unique(levels(data$forager_color))
```

Okay - some "red " with trailing space. Change these to "red" by removing trailing spaces from all levels.

```{r}
data$forager_color <- str_trim(data$forager_color)
unique(data$forager_color)
```

Assign treatment to colors for each colony.

```{r treatment}
# if colony == 1 & color == blue trt <- control
test <- vector(length = nrow(data))

data$trt <- ifelse(data$colony == 1 & data$forager_color == "blue", "control", 
                    ifelse(data$colony == 1 & data$forager_color == "red", "heat", 
                           ifelse(data$colony == 2 & data$forager_color == "blue", "control", 
                                  ifelse(data$colony == 2 & data$forager_color == "green", "heat", 
                                         NA))))
# check
table(data$forager_color, data$trt, data$colony)
```

For each colony, chi-squared test to evaluate if the proportion of ants foraging depends on heat treatment

```{r colony1}
data$trt <- as.factor(data$trt)
str(data)

data1 <- data[which(data$colony == 1 & !is.na(data$trt)), ]
str(data1)  

chisq.test(as.numeric(data1$trt))
plot(data1$trt)
```

No difference in foraging among treatments - almost exactly 50:50.

Same analyis for colony 2.

```{r colony2}
data2 <- data[which(data$colony == 2 & !is.na(data$trt)), ]
str(data2)  

chisq.test(as.numeric(data2$trt))
plot(data2$trt)
```

